
---
title: "Assignment QRM"
output: html_document
date: "2026-03-09"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(patchwork)
devtools::load_all()
```


# Start of Assignment

## Fetch data

```{r, include = FALSE}

data <- get_data()

```


# 1. Propose marginal distribution functions for each of the four stocks

We inspect each of the distributions by looking at the marginal distributions.

```{r, results =  'hide', echo = FALSE}
par(mfrow = c(2, 2))
fitdistrplus::descdist(data$GOOG, boot = 100)
fitdistrplus::descdist(data$MSFT, boot = 100)
fitdistrplus::descdist(data$MRK, boot = 100)
fitdistrplus::descdist(data$IDU, boot = 100)
par(mfrow = c(1, 1))

```

Here we have plottet respectively from top left to bottom right, GOOG, MSFT, MKR, IDU.

We have used the package **fitdistrplus** to plot the kurtosis vs the square of skewness. This should be a nice way to characterise the distributions. It measures the probability of extreme observations. 

$$
\begin{align*}
\frac{
    E[(X - \mu)^4]
}{
    \left( E[X - \mu)^2] \right)^2
}
\end{align*}
$$
We note that our distributions has a much larger kurtosis than that of a log-normal distribution.


Next to look at the asymmetry of the distributions, we look at the skewness.

$$
\begin{align*}
\frac{
    E[(X - \mu)^3]
}{
    \left( E[X - \mu)^2] \right)^{3/2}
}
\end{align*}
$$
The are all around zero, which indicates that the distributions are quite symmetric. 

Hence we will look at the t-distributions, since it can be parameterised with fatter tails, and still is symmetric. 


```{r, results =  'hide', echo = FALSE, warning = "FALSE", message = "FALSE"}

fit_goog <- t.dist.plot(ticker = "GOOG")
fit_msft <- t.dist.plot(ticker = "MSFT")
fit_mkr <- t.dist.plot(ticker = "MRK")
fit_idu <- t.dist.plot(ticker = "IDU")

(fit_goog$plot + fit_msft$plot) / (fit_mkr$plot + fit_idu$plot)

```